---
title: "simpleRegress"
author: "Hailey Watkins-Oliver"
date: "October 20, 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK 

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

## Model WEIGHT ~ AGE  this is an example  
```{r}
agemodel <- lm(WEIGHT~AGE,data=bears)
plot(WEIGHT~AGE,data=bears)
abline(agemodel)
summary.lm(agemodel)
```

1.Neck--> ARsqrd:0.87, RST:43.91
2.Age --> ARsqrd:0.55, RST:81.47

Plots on the simp graph are closer to the line than those on the agemodel graph, which displays its points far from it.

## Model WEIGHT ~ HEADLEN
```{r}
headlenmodel <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(headlenmodel)
summary.lm(headlenmodel)
```
Best to worst
1. Neck       -->     ARsqrd:0.87,  RST:43.91
2. Head length-->     ARsqrd:0.69, RST:67.81
3. Age        -->     ARsrd: 0.55,  RST:81.47

The headlenmodel graph is superior to the agemodel graph, despite its dispersed plots. When the agemodel's adjusted R-squared is 0.55 and the adjusted R-squared is 0.69, it is better than the simps' adjusted R-squared of 0.87, but still not good enough.

## Model WEIGHT ~ HEADWTH

```{r}
headwthmodel <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(headwthmodel)
summary.lm(headwthmodel)
```
Best to worst
1. Neck       -->   ARsqrd:0.87  RST:43.91
2. Head length-->   ARsqrd:0.69  RST:67.81
3. Head width -->   ARsqrd:0.60  RST:76.41
4. Age        -->   ARsrd: 0.55  RST:81.47

headwthmodel would have a high Resdiual standard error of 76.41, being higher than agemodel's 81.47 but not low enough for headlenmodel's 67.81 Residual standard error.